Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-20 analysis/paper/paper.Rmd
update plot and some spatial analysis

@@

```{r read-in-the-data}
library(here)
kwl_upper <- readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
```

```{r tidy-data, eval = TRUE}
# summarise the ornaments by layers
library(tidyverse)
library(ggplot2)

# L5 and L6 represent pre-European contact period, 
# L4 was the time of European contact, and 
# L2 and L1 was the period of Chinese contact.

@@

agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Agate bead") +
  theme_minimal(base_size = 14)
```

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring',
         !Type %in% c('large thick string',
                       'small thin string'))

ring_shape_plot <-
ggplot(ring_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Metal ring") +
  theme_minimal(base_size = 14)
```

@@

bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Bell") +
  theme_minimal(base_size = 14)
```

@@

glass_bead_plot <-
ggplot(glass_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Glass Bead") +
  theme_minimal(base_size = 14)
```

```{r plot-altogether, fig.cap= "Frequency of "}
library(cowplot)
plot_grid(agate_bead_shape_plot, ring_shape_plot,
          bell_shape_plot, glass_bead_plot,
          labels = c("A", "B", "C", "D"), ncol=2)
```
# Results

@@

## Frequency for spatial pattern
Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period. It shows that one square unearthed more ornaments than others in the pre-European contact. During the European contact period, the distribution is more ubiquitous with some concentration on northern part, and more even during Chinese contact period. Further look at the distribution for each major category individually, it shows that some squares have higher number of ornaments during the European contact period, including golden bead, agate bead, metal ring, and bell. However, there is no consistent pattern of concentration for different ornaments across the area. The squares with high number of ornaments distributed separately and independently without clear correlation. For example, more golden beads were found at the northern part, while more agate beads concentrated in the middle part.

Golden beads and

@@

>>>>

```{r read-in-the-data}
library(here)
library(tidyverse)
library(ggplot2)
kwl_upper <- readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
```

```{r tidy-data, eval = TRUE}
# summarise the ornaments by layers
# L5 and L6 represent pre-European contact period, 
# L4 was the time of European contact, and 
# L2 and L1 was the period of Chinese contact.

@@

@@

agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Agate bead") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring',
         !Type %in% c('large thick string',
                       'small thin string')) %>% 
  ungroup() %>% 
  mutate(Type = fct_drop(Type)) %>% 
  mutate(Type = fct_infreq(Type)) %>% 
  mutate(Type = fct_rev(Type))

ring_shape_plot <-
ggplot(ring_shape) +
  geom_col(aes(Type, n)) +
  coord_flip() +
  labs(y = "",
       x = "Metal ring") +
  facet_wrap(~ period) +
  theme_minimal(base_size = 14) 
```

@@

bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Bell") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

@@

glass_bead_plot <-
ggplot(glass_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Glass bead") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r plot-altogether, fig.cap= "Frequency of "}
library(cowplot)
plot_grid(ring_shape_plot,
          agate_bead_shape_plot, 
          bell_shape_plot, 
          glass_bead_plot,
          ncol=1,
          align = "h",
          axis = "lr",
          rel_widths = 1,
          rel_heights = c(1,1,0.7,0.62))
```
# Results

@@

@@

## Frequency for spatial pattern
Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all classes of ornaments for each time period. It shows that one square unearthed more ornaments than others in the pre-European contact. During the European contact period, the presence of ornaments is more ubiquitous with some clusters on northern part, and more randomly distributed during the Chinese contact period. When looking at the distribution for each major class individually, it shows that some clusters of different class of ornaments across the area during the European contact period, including golden bead, agate bead, metal ring, and bell. However, it seems no clear consistent pattern of concentration for those ornaments. Each class shows its own pattern that squares with higher number of ornaments distributed separately and independently. For example, the cluster of golden bead were found at the northern part, while the cluster of agate bead were found in the middle part. In contrast, there are more than one cluster of metal rings that distribute separately across the research area. Copper bells were usually found solely and seem randomly across the area. Before European contact, the map shows greater amount of glass beads were found at the northern and middle part of the research area, and more widespread during the European contact period. In the Chinese Contact period, both the amount and density of different classes of ornaments decreases that they were only found in a few squares, for example, golden bead and glass bead were only found in a couple squares at the middle part.

```{r point-pattern-data-setup}
# get centroid 
ornaments_period_join_cen <-
  ornaments_period_join %>% 
  mutate(lon = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
         lat = map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))

# plot the centroid
ggplot() +
  geom_sf(data = sample_units, fill = NA) +
  geom_point(data = ornaments_period_join_cen, 
             aes(x = lon, y = lat))

# transform data to PPP format and sp SpatialPoints
library(spatstat)

#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points
lon <- ornaments_period_join_cen$lon
lat <- ornaments_period_join_cen$lat
xrange <- range(lon, na.rm=T)
yrange <- range(lat, na.rm=T)

# create ppp
ornaments_point <- ppp(lon, lat, xrange, yrange)

To access whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.
